Supplementary information: In-silico molecular enrichment and clearance of the human intracranial space

### 2025-06-20 paper/supplementary.tex
Updates from Overleaf

\subsection{Nomenclature and notation} We begin by including \Cref{table:geometry_notation}, \Cref{table:variable}, and \Cref{table:mesh}, which summarize the common notation used for the geometry, model variables and parameters, and finite element discretization, respectively.
\begin{table}[h!] 
\centering
\caption{Geometrical notation, refer to (Fig.~1A, Fig.~2A--C)}
\label{table:geometry_notation}
\begin{tabular}{ll}
\hline
\textbf{Symbol} & \textbf{Description} \\
\hline
$\Omega$ & Full intracranial domain: $\Omega = \Omega_{\mathrm{PAR}} \cup \Omega_{\mathrm{CSF}}$ \\
$\Omega_{\mathrm{PAR}}$ & Brain parenchyma domain \\
$\Omega_{\mathrm{CSF}}$ & Cerebrospinal‐fluid (CSF) domain \\
$\partial\Omega_{\mathrm{PAR}}\cap\partial\Omega_{\mathrm{CSF}}$ & Parenchyma–CSF interface: $\partial \Omega_{\rm CSF} \bigcap \partial
\Omega_{\rm PAR} = \Gamma_{\rm LV} \cup \Gamma_{\rm pia}$ \\
$\Gamma_{\mathrm{LV}}$ & Lateral‐ventricle surface (part of parenchyma–CSF interface) \\
$\Gamma_{\mathrm{pia}}$ & Pial surface (remainder of parenchyma–CSF interface) \\
$\Gamma_{\mathrm{SSAS}}$ & Lower CSF boundary toward spinal subarachnoid space \\
$\Gamma_{\mathrm{AM}}$ & Outer CSF boundary toward arachnoid/dura \\
$\Gamma_{\mathrm{AM\text{-}L}}$, $\Gamma_{\mathrm{AM\text{-}U}}$ & Lower/upper parts of $\Gamma_{\mathrm{AM}}$ \\
$\Gamma_{\mathrm{SC}}$ & Parenchyma–spinal‐cord boundary: $\partial\Omega_{\mathrm{PAR}}\setminus\partial\Omega_{\mathrm{CSF}}$ \\
$\Lambda_{a},\Lambda_{v}$ & Arterial/venous PVS centerline networks \\
$\Lambda_{i}$ & Individual PVS centerline segment ($\Lambda_a^i$, $\Lambda_v^i$ corresponding to arterial and venous segments) \\
$\Omega_{i}$ & Perivascular space around $\Lambda_{i}$ \\
$R_{1}^{i},\,R_{2}^{i}$ & Inner/outer radii of the annular PVS $\Omega_{i}$ \\
$\Gamma_{a},\,\Gamma_{v}$ & Outer lateral surfaces of periarterial/perivenous PVSs \\
$P$ & Perimeter of the PVS cross‐section \\
$\mathcal V$ & Graph nodes of the PVS network \\
$\mathcal B\subset  \mathcal V$ & Internal (bifurcation) nodes \\
$\mathcal N\subset  \mathcal V$ & End (leaf) nodes \\
\hline
\end{tabular}
\end{table}

\begin{table}[htb!]
\centering 
\caption{Variable and parameter notation}
\label{table:variable}
\begin{tabular}{ll}
\hline
\textbf{Symbol} & \textbf{Description} \\
\hline
$\phi$ & Fluid volume fraction (porosity) \\
$D$ & Effective diffusion coefficient ($D_{\mathrm{PAR}} = D^*$ or $D_{\mathrm{CSF}} = D(1+R)$ (see eq. (1)) \\
$\mathbf{u}$ & Convective velocity ($\mathbf{u}_{\mathrm{PAR}} = 0$ or $\mathbf{u}_{\mathrm{CSF}}$ computed from steady state flow) \\
$c$ & Solute concentration in 3D domains \\
$\bar c$ & Lateral average of $c$ on the PVS boundary \\
$\xi$ & Permeability coefficient between the periarterial and perivenous networks
and their surroundings\\
$\delta_{\Gamma}$ & Dirac term on the PVS outer surfaces \\
$\beta_{\mathrm{pia}}$ & Permeability of pial/ventricle interface \\
$\beta_{\mathrm{exit}}$ & Efflux rate on $\Gamma_{\mathrm{AM\text{-}U}}$ \\
$g_{\mathrm{influx}}$ & Influx on $\Gamma_{\mathrm{SSAS}}$ \\
$A(s)$ & Cross‐sectional area of PVS at $s$ \\
$\hat D$ & 1D effective diffusion in PVS \\
$\hat u$ & 1D convective velocity in PVS \\
$p$ & Pressure in CSF (Stokes) or PVS (network) \\
$q$ & Volumetric flux in PVS network \\
$\kappa$ & Hydraulic conductivity in annular PVS \\
$\mu$ & Dynamic viscosity of CSF \\
$R_{0}$ & Resistance parameter for CSF efflux \\
$u_{\mathrm{in}}$ & Imposed fluid influx on $\Gamma_{\mathrm{LV}}$ \\
$\omega,\;\rho$ & Angular frequency, CSF density (in dispersion estimates) \\
\hline
\end{tabular}
\end{table}

\begin{table}[h!]
\centering
\caption{Common mesh and finite‐element notation}
\label{table:mesh}
\begin{tabular}{ll}
\hline
\textbf{Symbol} & \textbf{Description} \\
\hline
$\mathcal T$ & Tetrahedral mesh of $\Omega$ \\
$\mathcal{T}_{\mathrm{PAR}},\,T_{\mathrm{CSF}}$ & Restrictions of $\mathcal T$ to $\Omega_{\mathrm{PAR}},\Omega_{\mathrm{CSF}}$ \\
$E$ & Tetrahedral cell in $\mathcal T$ \\
$\mathcal F_{i}$ & Interior facets of $\mathcal T$ \\
$\mathcal F_{i,\mathrm{PAR}},\, \mathcal F_{i,\mathrm{CSF}}$ & Interior facets in $\Omega_{\mathrm{PAR}}$, $\Omega_{\mathrm{CSF}}$ \\
$\mathcal F_{e}$ & Exterior facets of CSF mesh ($\mathcal F_{\mathrm{pia}}\cup \mathcal F_{\mathrm{LV}}\cup \mathcal F_{\mathrm{AM\text{-}L}}\cup \mathcal \mathcal F_{\mathrm{AM\text{-}U}}\cup \mathcal F_{\mathrm{SSAS}}$) \\
$\mathcal F _{\mathrm{pia}},\,\mathcal F_{\mathrm{LV}},\, \mathcal F_{\mathrm{AM\text{-}L}},\,\mathcal F_{\mathrm{AM\text{-}U}},\, \mathcal F_{\mathrm{SSAS}}$ & Facets on corresponding boundaries \\
$V_{h}$ & Discontinuous Galerkin space on $\Omega$ of degree $1$\\
$\widehat V_{h}$ & Continuous Galerkin space on $\Lambda$ of degree $1$.  \\
% $\widehat V_{h,0}$, $\widehat V_{h,g}$ & Continuous Galerkin space on network with homogeneous or prescribed BCs \\
$\mathit{BDM}_{2}$ & Brezzi–Douglas–Marini space of degree 2 \\
$Q_{h}$ & Pressure space on $\mathcal T_{\mathrm{CSF}}$  \\
$I_{a},I_{v}$ & 1D meshes of arterial/venous networks $\Lambda_{a},\Lambda_{v}$ \\
$\widehat V_{h}(I)$ & Space of piecewise constants ($P^0$) on network mesh $I$ \\
\hline
\end{tabular}
\end{table}

\subsection{Flow and mixed-dimensional models: governing equations and corresponding finite element methods }
\label{sec:app:more_maths}

\subsubsection{Notation pertaining to the geometry and variables}

@@

\subsubsection{Computational mesh entities: definitions and common notation}

@@

\subsubsection{Modelling and simulating steady state CSF flow via the incompressible Stokes equations}
\label{sec:app:csf_flow}
The time-independent, incompressible Stokes equations model the flow
of an incompressible Newtonian fluid at low Reynolds numbers, and read
as follows: over the domain $\Omega_{\rm CSF} \subset \R^3$, find the
velocity vector field $\bm u : \Omega_{\rm CSF} \rightarrow \R^3$ and
the pressure field $p : \Omega_{\rm CSF} \rightarrow \R$ such that
\begin{subequations}
  \begin{alignat}{2}
    - \mu \Delta \bm u  + \nabla p &=  0 \quad && \text{ in } \Omega_{\rm CSF},
    \label{eq:momentum_equation} \\ 
    \nabla \cdot \bm u &= 0 \quad && \text{ in } \Omega_{\rm CSF}
    \label{eq:divergence_equation} .
  \end{alignat}
  \label{eq:stokes}%
\end{subequations}%
In addition, we impose the following boundary conditions to model flow
induced by CSF production in the choroid plexus with $\Gamma_{\rm AM-U}$ as the main efflux site:
\begin{subequations}
  \begin{alignat}{2}
    ( \mu \nabla \bm u \cdot \bm{n} - p \bm n) \cdot \bm n  &= -R_{\rm CSF, 0} ( \bm u \cdot \bm n ), 
    \quad \bm u \cdot \bm t = 0
    && \text{ on } \Gamma_{\mathrm{AM-U}}, \\
    \bm u &= 0 
    && \text{ on } \Gamma_{\mathrm{AM-L}}\cup\Gamma_{\rm{pia}}\cup\Gamma_{\rm{SSAS}}, \\
    %\bm u &= 0 && \text{ on } \Gamma_{\rm{pia}}, \\
    %\bm u &= 0 && \text{ on } \Gamma_{\rm{SSAS}}, \\
    \bm u \cdot \bm n & = \frac{1}{ |\Gamma_{\rm LV}|}  u_{\rm in}, 
    \quad \quad \bm u \cdot \bm t = 0 \quad 
    && \text{ on } \Gamma_{\rm{LV}} ,  
  \end{alignat}
  \label{eq:stokes_bcs}%
\end{subequations}%
where $\bm n$ and $\bm t$ denote the unit outward normal and tangent vectors to the boundary respectively,  $\mu$ is the (dynamic) CSF viscosity, $R_{\rm CSF, 0} > 0$ is a resistance parameter for CSF efflux, and $u_{\rm in}$ is a given fluid influx, here across the lateral ventricle wall $\Gamma_{\rm LV}$. We consider alternative variations of these boundary conditions in connection with estimating dispersion effects induced by CSF pulsatility, see~\Cref{sec:app:dispersion}.

@@

\subsubsection{Modeling and simulating steady flow in perivascular networks induced by pressure differences}
\label{sec:app:hydraulic_network}

@@

\textit{Finite element solution of the perivascular network equations}. 
We consider meshes $\mathcal{I}_a, \mathcal{I}_v$ representing a
conforming subdivision of each of the perivascular networks
$\Lambda_a$, $\Lambda_v$. Relative to each mesh $\mathcal{I}$, we define the
space of (discontinuous) piecewise constants $\hat{Z}_h(\mathcal{I})$ and define
spaces of continuous piecewise linears $\hat{V}_{h, g}(\mathcal{I})$ with
prescribed boundary node values (on $\mathcal{N}$) given by $g$:
\begin{align}
  \hat{Z}_h(\mathcal{I}) &= \{z \in L^2(\Lambda) ; \, z \vert_{\Lambda_i} \in \mathrm{P}^{0}(\Lambda_i) \, \foralls \Lambda^i \in \mathcal{I}\}, \\
  \hat{V}_{h, g}(\mathcal{I}) &= \{ v \in C^0(\Lambda) ; \, v \vert_{\Lambda^i} \in \mathrm{P}^1(\Lambda_i), \, v(x) = g(x) \, \foralls x \in \mathcal{N} \} .
  %Q_h^{\rm PVS} &= \{ q_h \in C^0(\Lambda),  q_h \vert_{\Lambda^i} \in \mathrm{P}^k(\Lambda_i), q_h \vert_{\Gamma_{\rm Pia}} = p_{\rm CSF} \}
\end{align}
To discretize~\eqref{eq:1d_flow} with the bifurcation
conditions~\eqref{eq:1d_flow:bifurcations} and boundary
conditions~\eqref{eq:1d_flow:bcs}, we use the space of continuous
functions $\hat{Z}_h(\mathcal{I})$ to enforce the continuity of the pressure at
bifurcation points, while the conservation of flux is enforced
(weakly) through the variational formulation. The discrete variational form of the equations then reads: Find $(\hat{q}, \hat{p}) \in \hat{Z}_h(\mathcal{I}) \times \hat{V}_{h, p_0}(\mathcal{I})$ such that
\begin{subequations}
\begin{align}
  \langle \hat{q}, z \rangle +  \langle z, \frac{A \kappa}{\mu} \partial_s \hat{p}  \rangle &= 0 \quad \forall z \in \hat{Z}_h(\mathcal{I}) \\
  \langle \hat{q}, \partial_s w \rangle &= 0 \quad  \forall w \in \hat{V}_{h, 0}(\mathcal{I}).
\end{align}
\end{subequations}%
where $\langle \cdot, \cdot \rangle$ denotes the $L^2(\mathcal{I})$-inner product and is defined segment-wise. For a stability and convergence analysis of this discrete model, we refer to the reference\cite{gjerde2024directional}.

\subsubsection{Modeling and simulating solute transport and exchange in intracranial domains (3D) and perivascular networks (1D)}
\label{sec:app:transport_eqs}

@@

@@

We use the theoretical framework previously introduced by Gjerde et al.~\cite{gjerde2023directional} to compute an analytic estimate of the time-averaged (or \emph{net}) flow rates $\langle Q'_i \rangle$ (mm$^3$/s) induced by peristaltic pumping in a perivascular network $\Lambda = \cup_i \Lambda_i$. The motion of the (inner) vascular wall is assumed to be described by a periodic traveling (peristaltic) wave of relative amplitude $\varepsilon$, wave length $\lambda$ (mm) and frequency $f$, acting normal to the wall and traveling in the antegrade direction (with the blood flow). By definition, the wave number is $k = 2 \pi/\lambda$ and the angular frequency is $\omega = 2 \pi f$. Each PVS segment $\Lambda_i$ has length $L_i$ with wave-relative length $\ell_i = k L_i$, baseline inner radius $R_1^i$, fixed outer radius $R_2^i$, and outer-to-inner ratio $\beta_i = R_2^i/R_1^i$. These geometric parameters and the assumption of annular cylindrical PVS segments yield hydraulic resistances $\mathcal{R}_{o, i}$ and additional characteristic parameters, see \cite{gjerde2023directional} for the complete definitions and schematics. Since the analytical estimate is derived under the assumption that $k L_i \approx \mathcal{O} (1)$~\cite{gjerde2023directional} and has been verified against numerical simulations for $k L_i$ of the order $10^{-1}-10^2$~\cite[Table I]{gjerde2023directional}, we consider it applicable for the wave lengths and vascular network data considered here in which $k L_i$ range from 0.15 to 30 for the strong vasomotion and 0.0015 to 0.30 for the cardiac waves (\Cref{tab:pvs:parameters}).
\begin{table}
\centering
  \small
  \begin{tabular}{llrc}
    \toprule
    Parameter & Description & Value(s)  & Ref.\\ 
    \midrule
    Relative PVS size $\beta$ & $\beta = \beta_i = R_2^i / R_1^i$ & 2 & $\approx$\cite{eide2024functional} \\
    Wave frequency $f$ & Traveling wave frequency of vascular motion & $0.1-1.0$ Hz & \cite{gjerde2023directional} \\
    Wave length $\lambda$ & Traveling wave length of vascular motion & $0.02-2$ mm & \cite{broggini2024long, gjerde2023directional} \\
    Wave amplitude & Relative amplitude of inner wall motion & $1-10\%$ & \cite{gjerde2023directional} \\
    \midrule
    Cardiac frequency $f_c$ & Frequency of human cardiac pulse wave  & $1.0$ Hz & \cite{jung2021novel} \\
    Cardiac wave length $\lambda_c$ & Wave length of human cardiac pulse wave & $2.0$m & \\
    Cardiac amplitude $\varepsilon_c$ & Wall displacement amplitude of cardiac pulse wave & 1\%  & \\
    \midrule
    Vasomotion frequency $f_v$ & Frequency of slow vasomotion wave  & $0.1$ Hz & \cite{jung2021novel} \\
    Vasomotion wave length $\lambda_v$ & Wave length of vasomotion wave & $0.02$ m & \cite{broggini2024long, gjerde2023directional} \\
    Vasomotion amplitude $\varepsilon_v$ & Wall displacement amplitude of vasomotion wave & 10\%  & \cite{broggini2024long, gjerde2023directional}\\
    \bottomrule
  \end{tabular}
  \caption{Perivascular flow induced by vascular wall motion: overview of parameters.}
  \label{tab:pvs:parameters}
\end{table}
% Vascular radii r_o (min, max, avg, std) (mm):  ....
% Vascular branch lengths L (avg, std, min, max) (mm): ...

>>>>

\subsection{Nomenclature and notation} We begin by including \Cref{table:geometry_notation}, \Cref{table:variable}, and \Cref{table:mesh}, which summarize the common notation used for the geometry, model variables and parameters, and finite element discretization, respectively.
\begin{table}[h!] 
\centering
\caption{Geometrical notation, refer to (Fig.~1A, Fig.~2A--C)}
\label{table:geometry_notation}
\begin{tabular}{ll}
\hline
\textbf{Symbol} & \textbf{Description} \\
\hline
$\Omega$ & Full intracranial domain: $\Omega = \Omega_{\mathrm{PAR}} \cup \Omega_{\mathrm{CSF}}$ \\
$\Omega_{\mathrm{PAR}}$ & Brain parenchyma domain \\
$\Omega_{\mathrm{CSF}}$ & Cerebrospinal‐fluid (CSF) domain \\
$\partial\Omega_{\mathrm{PAR}}\cap\partial\Omega_{\mathrm{CSF}}$ & Parenchyma–CSF interface: $\partial \Omega_{\rm CSF} \bigcap \partial
\Omega_{\rm PAR} = \Gamma_{\rm LV} \cup \Gamma_{\rm pia}$ \\
$\Gamma_{\mathrm{LV}}$ & Lateral‐ventricle surface (part of parenchyma–CSF interface) \\
$\Gamma_{\mathrm{pia}}$ & Pial surface (remainder of parenchyma–CSF interface) \\
$\Gamma_{\mathrm{SSAS}}$ & Lower CSF boundary toward spinal subarachnoid space \\
$\Gamma_{\mathrm{AM}}$ & Outer CSF boundary toward arachnoid/dura \\
$\Gamma_{\mathrm{AM\text{-}L}}$, $\Gamma_{\mathrm{AM\text{-}U}}$ & Lower/upper parts of $\Gamma_{\mathrm{AM}}$ \\
$\Gamma_{\mathrm{SC}}$ & Parenchyma–spinal‐cord boundary: $\partial\Omega_{\mathrm{PAR}}\setminus\partial\Omega_{\mathrm{CSF}}$ \\
$\Lambda_{a},\Lambda_{v}$ & Arterial/venous PVS centerline networks \\
$\Lambda_{i}$ & Individual PVS centerline segment ($\Lambda_a^i$, $\Lambda_v^i$ corresponding to arterial and venous segments) \\
$L_i$ & Length of a PVS segment $\Lambda_i$ \\
$\Omega_{i}$ & Perivascular space around $\Lambda_{i}$ \\
$R_{1}^{i},\,R_{2}^{i}$ & Inner/outer radii of the annular PVS $\Omega_{i}$ \\
$\Gamma_{a},\,\Gamma_{v}$ & Outer lateral surfaces of periarterial/perivenous PVSs \\
$P$ & Perimeter of the PVS cross‐section \\
$\mathcal V$ & Graph nodes of the PVS network \\
$\mathcal B\subset  \mathcal V$ & Internal (bifurcation) nodes \\
$\mathcal N\subset  \mathcal V$ & End (leaf) nodes \\
\hline
\end{tabular}
\end{table}

\begin{table}[htb!]
\centering 
\caption{Variable and parameter notation}
\label{table:variable}
\begin{tabular}{ll}
\hline
\textbf{Symbol} & \textbf{Description} \\
\hline
$\phi$ & Fluid volume fraction (porosity) \\
$D$ & Effective diffusion coefficient ($D_{\mathrm{PAR}} = D^*$ or $D_{\mathrm{CSF}} = D(1+R)$ \\
$\mathbf{u}$ & Convective velocity ($\mathbf{u}_{\mathrm{PAR}} = 0$ or $\mathbf{u}_{\mathrm{CSF}}$ computed from steady state flow) \\
$c$ & Solute concentration in 3D domains \\
$\bar c$ & Lateral average of $c$ on the PVS boundary \\
$\xi$ & Permeability coefficient between the periarterial and perivenous networks
and their surroundings\\
$\delta_{\Gamma}$ & Dirac term on the PVS outer surfaces \\
$\beta_{\mathrm{pia}}$ & Permeability of pial/ventricle interface \\
$\beta_{\mathrm{exit}}$ & Efflux rate on $\Gamma_{\mathrm{AM\text{-}U}}$ \\
$g_{\mathrm{influx}}$ & Influx on $\Gamma_{\mathrm{SSAS}}$ \\
$A(s)$ & Cross‐sectional area of PVS at $s$ \\
$\hat D$ & 1D effective diffusion in PVS \\
$\hat u$ & 1D convective velocity in PVS \\
$p$ & Pressure in CSF (Stokes) or PVS (network) \\
$\hat{q}$ & Volumetric flux in PVS network \\
$\kappa$ & Hydraulic conductivity in annular PVS \\
$\mu$ & Dynamic viscosity of CSF \\
$R_{\rm CSF, 0}$ & Resistance parameter for CSF efflux \\
$u_{\mathrm{in}}$ & Imposed fluid influx on $\Gamma_{\mathrm{LV}}$ \\
$\omega,\;\rho$ & Angular frequency, CSF density (for dispersion estimates) \\
$\alpha_c,\, \alpha_r$ & Womersley number (cardiac, respiratory) \\
$h_{\rm SAS}$ & Mean width of the subarachnoid space (for dispersion estimates) \\
$R_c,\,R_r$ & Spatially-varying dispersion enhancement factors (cardiac, respiratory) \\

\hline
\end{tabular}
\end{table}

\begin{table}[h!]
\centering
\caption{Common mesh and finite‐element notation}
\label{table:mesh}
\begin{tabular}{ll}
\hline
\textbf{Symbol} & \textbf{Description} \\
\hline
$\mathcal{T}$ & Tetrahedral mesh of $\Omega$ \\
$\mathcal{T}_{\mathrm{PAR}},\,\mathcal{T}_{\mathrm{CSF}}$ & Restrictions of $\mathcal T$ to $\Omega_{\mathrm{PAR}},\Omega_{\mathrm{CSF}}$ \\
$E$ & Tetrahedral cell in $\mathcal T$ \\
$\mathcal F_{i}$ & Interior facets of $\mathcal T$ \\
$\mathcal F_{i,\mathrm{PAR}},\, \mathcal F_{i,\mathrm{CSF}}$ & Interior facets in $\Omega_{\mathrm{PAR}}$, $\Omega_{\mathrm{CSF}}$ \\
$\mathcal F_{e}$ & Exterior facets of CSF mesh ($\mathcal F_{\mathrm{pia}}\cup \mathcal F_{\mathrm{LV}}\cup \mathcal F_{\mathrm{AM\text{-}L}}\cup \mathcal \mathcal F_{\mathrm{AM\text{-}U}}\cup \mathcal F_{\mathrm{SSAS}}$) \\
$\mathcal F _{\mathrm{pia}},\,\mathcal F_{\mathrm{LV}},\, \mathcal F_{\mathrm{AM\text{-}L}},\,\mathcal F_{\mathrm{AM\text{-}U}},\, \mathcal F_{\mathrm{SSAS}}$ & Facets on corresponding boundaries \\
$V_{h}$ & Discontinuous Galerkin space on $\Omega$ of degree $1$\\
$\widehat V_{h}$ & Continuous Galerkin space on $\Lambda$ of degree $1$.  \\
% $\widehat V_{h,0}$, $\widehat V_{h,g}$ & Continuous Galerkin space on network with homogeneous or prescribed BCs \\
$\mathrm{BDM}_{2}$ & Brezzi–Douglas–Marini space of degree 2 \\
$Q_{h}$ & Pressure space on $\mathcal T_{\mathrm{CSF}}$  \\
$\mathcal{I}_a, \mathcal{I}_v$ & 1D meshes of arterial/venous networks $\Lambda_{a},\Lambda_{v}$ \\
$\widehat{Z}_{h}(\mathcal{I})$ & Space of piecewise constants ($P^0$) on network mesh $I$ \\
$\tau$ & Time step size \\
\hline
\end{tabular}
\end{table}

\subsection{Notation pertaining to the geometry, variables and computational domains}

@@

% \subsection{Computational mesh entities: definitions and common notation}

@@

%\subsection{Flow and mixed-dimensional models: governing equations and corresponding finite element methods }
%\label{sec:app:more_maths} 
\subsection{Modelling and simulating steady state CSF flow via the incompressible Stokes equations}
\label{sec:app:csf_flow}
The time-independent, incompressible Stokes equations model the flow
of an incompressible Newtonian fluid at low Reynolds numbers, and read
as follows: over the domain $\Omega_{\rm CSF} \subset \R^3$, find the
velocity vector field $\bm u : \Omega_{\rm CSF} \rightarrow \R^3$ and
the pressure field $p : \Omega_{\rm CSF} \rightarrow \R$ such that
\begin{subequations}
  \begin{alignat}{2}
    - \mu \Delta \bm u  + \nabla p &=  0 \quad && \text{ in } \Omega_{\rm CSF},
    \label{eq:momentum_equation} \\ 
    \nabla \cdot \bm u &= 0 \quad && \text{ in } \Omega_{\rm CSF}
    \label{eq:divergence_equation} .
  \end{alignat}
  \label{eq:stokes}%
\end{subequations}%
In addition, we impose the following boundary conditions to model flow
induced by CSF production in the choroid plexus with $\Gamma_{\rm AM-U}$ as the main efflux site:
\begin{subequations}
  \begin{alignat}{2}
    ( \mu \nabla \bm u \cdot \bm{n} - p \bm n) \cdot \bm n  &= -R_{\rm CSF, 0} ( \bm u \cdot \bm n ), 
    \quad \bm u \cdot \bm t = 0
    && \text{ on } \Gamma_{\mathrm{AM-U}}, \\
    \bm u &= 0 
    && \text{ on } \Gamma_{\mathrm{AM-L}}\cup\Gamma_{\rm{pia}}\cup\Gamma_{\rm{SSAS}}, \\
    %\bm u &= 0 && \text{ on } \Gamma_{\rm{pia}}, \\
    %\bm u &= 0 && \text{ on } \Gamma_{\rm{SSAS}}, \\
    \bm u \cdot \bm n & = \frac{1}{ |\Gamma_{\rm LV}|}  u_{\rm in}, 
    \quad \quad \bm u \cdot \bm t = 0 \quad 
    && \text{ on } \Gamma_{\rm{LV}} ,  
  \end{alignat}
  \label{eq:stokes_bcs}%
\end{subequations}%
where $\bm n$ and $\bm t$ denote the unit outward normal and tangent vectors to the boundary respectively,  $\mu$ is the (dynamic) CSF viscosity, $R_{\rm CSF, 0} > 0$ is a resistance parameter for CSF efflux, and $u_{\rm in}$ is a given fluid influx, here across the lateral ventricle wall $\Gamma_{\rm LV}$. We consider alternative variations of these boundary conditions in connection with estimating dispersion effects induced by CSF pulsatility, see~\Cref{sec:app:dispersion}.

@@

\subsection{Modeling and simulating steady flow in perivascular networks induced by pressure differences}
\label{sec:app:hydraulic_network}

@@

\textit{Finite element solution of the perivascular network equations}. 
We consider meshes $\mathcal{I}_a, \mathcal{I}_v$ representing a
conforming subdivision of each of the perivascular networks
$\Lambda_a$, $\Lambda_v$. Relative to each mesh $\mathcal{I}$, we define the
space of (discontinuous) piecewise constants $\widehat{Z}_h(\mathcal{I})$ and define
spaces of continuous piecewise linears $\hat{V}_{h, g}(\mathcal{I})$ with
prescribed boundary node values (on $\mathcal{N}$) given by $g$:
\begin{align}
  \hat{Z}_h(\mathcal{I}) &= \{z \in L^2(\Lambda) ; \, z \vert_{\Lambda_i} \in \mathrm{P}^{0}(\Lambda_i) \, \foralls \Lambda^i \in \mathcal{I}\}, \\
  \hat{V}_{h, g}(\mathcal{I}) &= \{ v \in C^0(\Lambda) ; \, v \vert_{\Lambda^i} \in \mathrm{P}^1(\Lambda_i), \, v(x) = g(x) \, \foralls x \in \mathcal{N} \} .
  %Q_h^{\rm PVS} &= \{ q_h \in C^0(\Lambda),  q_h \vert_{\Lambda^i} \in \mathrm{P}^k(\Lambda_i), q_h \vert_{\Gamma_{\rm Pia}} = p_{\rm CSF} \}
\end{align}
To discretize~\eqref{eq:1d_flow} with the bifurcation
conditions~\eqref{eq:1d_flow:bifurcations} and boundary
conditions~\eqref{eq:1d_flow:bcs}, we use the space of continuous
functions $\hat{Z}_h(\mathcal{I})$ to enforce the continuity of the pressure at
bifurcation points, while the conservation of flux is enforced
(weakly) through the variational formulation. The discrete variational form of the equations then reads: Find $(\hat{q}, \hat{p}) \in \hat{Z}_h(\mathcal{I}) \times \hat{V}_{h, p_0}(\mathcal{I})$ such that
\begin{subequations}
\begin{align}
  \langle \hat{q}, z \rangle +  \langle z, \frac{A \kappa}{\mu} \partial_s \hat{p}  \rangle &= 0 \quad \forall z \in \hat{Z}_h(\mathcal{I}) \\
  \langle \hat{q}, \partial_s w \rangle &= 0 \quad  \forall w \in \hat{V}_{h, 0}(\mathcal{I}).
\end{align}
\end{subequations}%
where $\langle \cdot, \cdot \rangle$ denotes the $L^2(\mathcal{I})$-inner product and is defined segment-wise. For a stability and convergence analysis of this discrete model, we refer to the reference\cite{gjerde2024directional}.

@@

We use the theoretical framework previously introduced by Gjerde et al.~\cite{gjerde2023directional} to compute an analytic estimate of the time-averaged (or \emph{net}) flow rates $\langle Q'_i \rangle$ (mm$^3$/s) induced by peristaltic pumping in a perivascular network $\Lambda = \cup_i \Lambda_i$. The motion of the (inner) vascular wall is assumed to be described by a periodic traveling (peristaltic) wave of relative amplitude $\varepsilon$, wave length $\lambda$ (mm) and frequency $f$, acting normal to the wall and traveling in the antegrade direction (with the blood flow). By definition, the wave number is $k = 2 \pi/\lambda$ and the angular frequency is $\omega = 2 \pi f$. Each PVS segment $\Lambda_i$ has length $L_i$ with wave-relative length $\ell_i = k L_i$, baseline inner radius $R_1^i$, fixed outer radius $R_2^i$, and outer-to-inner ratio $\beta_i = R_2^i/R_1^i$. These geometric parameters and the assumption of annular cylindrical PVS segments yield hydraulic resistances and additional characteristic parameters, see \cite{gjerde2023directional} for the complete definitions and schematics. Since the analytical estimate is derived under the assumption that $k L_i \approx \mathcal{O} (1)$~\cite{gjerde2023directional} and has been verified against numerical simulations for $k L_i$ of the order $10^{-1}-10^2$~\cite[Table I]{gjerde2023directional}, we consider it applicable for the wave lengths and vascular network data considered here in which $k L_i$ range from 0.15 to 30 for the strong vasomotion and 0.0015 to 0.30 for the cardiac waves (\Cref{tab:pvs:parameters}).
\begin{table}
\centering
  \small
  \begin{tabular}{llrc}
    \toprule
    Parameter & Description & Value(s)  & Ref.\\ 
    \midrule
    Relative PVS size $\beta$ & $\beta = \beta_i = R_2^i / R_1^i$ & 2 & $\approx$\cite{eide2024functional} \\
    Wave frequency $f$ & Traveling wave frequency of vascular motion & $0.1-1.0$ Hz & \cite{gjerde2023directional} \\
    Wave length $\lambda$ & Traveling wave length of vascular motion & $0.02-2$ mm & \cite{broggini2024long, gjerde2023directional} \\
    Wave amplitude & Relative amplitude of inner wall motion & $1-10\%$ & \cite{gjerde2023directional} \\
    \midrule
    Cardiac frequency $f_c$ & Frequency of human cardiac pulse wave  & $1.0$ Hz & \cite{jung2021novel} \\
    Cardiac wave length $\lambda_c$ & Wave length of human cardiac pulse wave & $2.0$m & \\
    Cardiac amplitude $\varepsilon_c$ & Wall displacement amplitude of cardiac pulse wave & 1\%  & \\
    \midrule
    Vasomotion frequency $f_v$ & Frequency of slow vasomotion wave  & $0.1$ Hz & \cite{jung2021novel} \\
    Vasomotion wave length $\lambda_v$ & Wave length of vasomotion wave & $0.02$ m & \cite{broggini2024long, gjerde2023directional} \\
    Vasomotion amplitude $\varepsilon_v$ & Wall displacement amplitude of vasomotion wave & 10\%  & \cite{broggini2024long, gjerde2023directional}\\
    \bottomrule
  \end{tabular}
  \caption{Perivascular flow induced by vascular wall motion: overview of parameters.}
  \label{tab:pvs:parameters}
\end{table}
% Vascular radii r_o (min, max, avg, std) (mm):  ....
% Vascular branch lengths L (avg, std, min, max) (mm): ...

\subsection{Modeling and simulating solute transport and exchange in intracranial domains (3D) and perivascular networks (1D)}
\label{sec:app:transport_eqs}

@@
